---
title: "Expression vector graphs, stats, and data wrangling"
author: "Jennifer Tran"
date: "2024-05-13"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load necessary packages for these graphs:

```{r, warning=FALSE, message = FALSE}
require('pacman')

p_load(dplyr, data.table, ggplot2, tidyr, RColorBrewer, stringr, ggridges, colourpicker, ggbreak)
```

## Conjugative transfer and transposition efficency graphs (Fig S1)

#### Plasmid conjugation

Load plasmid data from spreadsheet (CFUs in LB vs LB+Kanamycin) as plasmid_eff

```{r echo = FALSE}
plasmid_eff <- fread("./20240426_plasmid_efficiencies.csv", header=TRUE)
```

```{r echo=FALSE}
print(plasmid_eff)
```

Calculate efficiency if not already included

```{r}
plasmid_eff$efficiency <- plasmid_eff$LB_Kan/plasmid_eff$LB
```

Filter for *A. baumannii* ATCC 17978 and *E. coli* BW25113 empty vector data  
Calculate mean efficiences for each

```{r}
plasmid_plot <- plasmid_eff %>%
  filter(parent %in% c("17978", "BW25113"), plasmid == "EV")

mean_efficiencies <- plasmid_plot %>%
  group_by(parent) %>%
  summarise(
    mean_efficiency = mean(efficiency),
    .groups = 'drop'
  )
```

```{r echo=FALSE}
print(mean_efficiencies)
```

Plot a dot plot with the mean as a bar and dashed line for limit of detection

```{r}
ggplot() +
  geom_errorbar(data = mean_efficiencies, 
                aes(x = parent, ymin = mean_efficiency, ymax = mean_efficiency),
                width = 0.2, linewidth = 1.5, color = "black") +
  geom_point(data = plasmid_plot, 
             aes(x = parent, y = efficiency, color = parent), 
             size = 4, alpha = 0.5) +
  geom_hline(yintercept = 0.33e-9, linetype = "dashed", color = "black") +
  scale_y_log10(
    limits = c(1e-10, 1.5),
    breaks = c(1e-10, 1e-8, 1e-6, 1e-4, 1e-2, 1),
    labels = c(expression(10^-10), expression(10^-8), expression(10^-6), 
               expression(10^-4), expression(10^-2), expression(10^0))) +
  labs(x = "Strain", y = "Conjugative Efficiency", 
       title = "Plasmid Transfer Efficiency") +
  theme_minimal() +
  theme(legend.position = "none")
```

#### Tn7 transposition

Load Tn7 data from spreadsheet (CFUs in LB vs LB+apramycin)

```{r, echo=FALSE}
Tn7_eff <- fread("./20240419_Tn7_efficiences.csv", header=TRUE)
```

```{r echo=FALSE}
print(Tn7_eff)
```

Calculate efficiency if not already included

```{r}
Tn7_eff$efficiency <- Tn7_eff$LB_Apr/Tn7_eff$LB
```

Filter for *A. baumannii* ATCC 17978 and *E. coli* BW25113 empty vector data\
Calculate the mean efficiencies for each

```{r}
Tn7_plot <- Tn7_eff %>%
  filter(parent %in% c("17978", "BW25113"), Tn7 == "EV")

mean_efficiencies <- Tn7_plot %>%
  group_by(parent) %>%
  summarise(
    mean_efficiency = mean(efficiency),
    .groups = 'drop'
  )
```

```{r echo=FALSE}
print(mean_efficiencies)
```

Plot a dot plot with the mean as a bar

```{r}
ggplot() +
  geom_errorbar(data = mean_efficiencies, 
                aes(x = parent, ymin = mean_efficiency, ymax = mean_efficiency),
                width = 0.2, linewidth = 1.5, color = "black") +
  geom_point(data = Tn7_plot, aes(x = parent, y = efficiency, color = parent), 
             size = 4, alpha = 0.5) +
  geom_hline(yintercept = (0.33e-10), linetype = "dashed", color = "black") +
  scale_y_log10(
    limits = c(1e-11, 1.5),
    breaks = c(1e-10, 1e-8, 1e-6, 1e-4, 1e-2, 1),
    labels = c(expression(10^-10), expression(10^-8), expression(10^-6), 
               expression(10^-4), expression(10^-2), expression(10^0))) +
  labs(x = "Strain", y = "Transposition Efficiency", 
       title = "Tn7 Transposition Efficiency") +
  theme_minimal() +
  theme(legend.position = "none")
```

## Promoter expression dot plots (Fig 2)

Load cleaned plate reader data

```{r echo=FALSE}
IPTG_data <- read.csv("20231122_17_gfp_promoters.csv", header = TRUE) %>%
  filter(dil != 0.2) %>% select(-c(dil))

print(IPTG_data)
```

Subtract background values from measurements

```{r}
columns_to_normalize <- c('Pabst.fluo', 'Pabst.OD', 'PabstBR.fluo', 
                          'PabstBR.OD', 'Ptrc.fluo', 'Ptrc.OD')

background_vals <- IPTG_data %>% filter(IPTG_conc %like% "background") %>% 
  select(all_of(columns_to_normalize))

for (col in columns_to_normalize) {
  IPTG_data[[col]] <- IPTG_data[[col]] - background_vals[[col]]
}

IPTG_data <- IPTG_data %>% filter(!(IPTG_conc %like% "background")) 
```

Normalize to cell density (fluorescence/OD)

```{r}
fluo_columns <- grep("\\.fluo$", names(IPTG_data), value = TRUE)
od_columns <- sub("fluo", "OD", fluo_columns)

for (i in seq_along(fluo_columns)) {
  new_col_name <- paste0(sub("\\.fluo", "", fluo_columns[i]), ".Ratio")
  IPTG_data[[new_col_name]] <- IPTG_data[[fluo_columns[i]]] / 
    IPTG_data[[od_columns[i]]]
}

data.norm <- IPTG_data %>%
  select(IPTG_conc, rep, ends_with(".Ratio"))
```

Subtract empty vector noise (autofluorescence)\
Additionally calculate SD with propagated error

```{r}
#empty vector stats
ev_stats <- data.norm %>%
  filter(IPTG_conc == "EV") %>%
  summarise(across(ends_with(".Ratio"), list(mean = mean, sd = sd)))

#sample stats
mean_sd_diff <- data.norm %>%
  filter(IPTG_conc != "EV") %>%
  group_by(IPTG_conc) %>%
  summarise(across(ends_with(".Ratio"), list(mean = mean, sd = sd)))

#subtract off EV
data_adjusted <- data.norm %>%
  filter(IPTG_conc != "EV") %>%
  rowwise() %>%
  mutate(
    Ptrc.Ratio = Ptrc.Ratio - ev_stats$Ptrc.Ratio_mean,
    Pabst.Ratio = Pabst.Ratio - ev_stats$Pabst.Ratio_mean,
    PabstBR.Ratio = PabstBR.Ratio - ev_stats$PabstBR.Ratio_mean
  )

#make data long format for ggplot downstream
data_adjusted_long <- data_adjusted %>%
  pivot_longer(
    cols = starts_with("p"),
    names_to = "promoter",
    values_to = "value"
)

mean_sd_adjusted <- data_adjusted %>%
  group_by(IPTG_conc) %>%
  summarise(across(ends_with(".Ratio"), list(mean = mean, sd = sd)))

#final stats (means and SD)
error_propagation <- mean_sd_adjusted %>%
  mutate(
    Ptrc.Ratio_PropagatedError = sqrt(mean_sd_diff$Ptrc.Ratio_sd^2 + 
                                        ev_stats$Ptrc.Ratio_sd^2),
    Pabst.Ratio_PropagatedError = sqrt(mean_sd_diff$Pabst.Ratio_sd^2 + 
                                         ev_stats$Pabst.Ratio_sd^2),
    PabstBR.Ratio_PropagatedError = sqrt(mean_sd_diff$PabstBR.Ratio_sd^2 + 
                                           ev_stats$PabstBR.Ratio_sd^2)
  ) %>%
  select(IPTG_conc, ends_with("_mean"), ends_with("PropagatedError")) %>%
  pivot_longer(
    cols = -IPTG_conc, 
    names_to = c("Sample", ".value"),
    names_pattern = "(.*)(mean|PropagatedError)$"
  ) %>%
  mutate(Sample = str_remove(Sample, "_"))

replicate_counts <- aggregate(rep ~ IPTG_conc, data = data.norm, FUN = length)
names(replicate_counts)[2] <- "Num_Replicates"

error_propagation <- merge(error_propagation, replicate_counts, 
                           by = "IPTG_conc", all.x = TRUE)
```

```{r echo=FALSE}
print(error_propagation)
```

Prepare data for plotting

```{r}
plot_data <- data_adjusted_long %>%
  group_by(IPTG_conc, promoter) %>%
  summarize(mean_value = mean(value))
```

#### Plots

```{r}
ggplot() +
  geom_errorbar(data = plot_data, aes(x = as.factor(promoter), 
                                      ymin = mean_value, ymax = mean_value),
                width = 0.4, size = 2, color = "black") +
  geom_point(data = data_adjusted_long, 
             aes(x = as.factor(promoter), y = value), size = 4, alpha = 0.5) +
  facet_grid(. ~ IPTG_conc) +
  labs(x = "IPTG Concentration", 
       y = "Relative Expression (normalized fluorescence)", 
       title = "Expression Levels for Different Samples 
       at IPTG Concentrations of 0 and 1mM") +
  theme_minimal() +
  theme(legend.position = "none") +
  scale_y_log10(limits = c(1, NA))
```

Calculate statistics using propagated error as SD

```{r}
# Separate the data by IPTG concentration
data_0 <- subset(error_propagation, IPTG_conc == 0)
data_1000 <- subset(error_propagation, IPTG_conc == 1000)

# Function to manually perform pairwise Welch's t-test
pairwise_t_test <- function(data) {
  combinations <- combn(unique(data$Sample), 2)
  results <- apply(combinations, 2, function(combo) {
    sample1 <- data[data$Sample == combo[1],]
    sample2 <- data[data$Sample == combo[2],]
    
    # Calculate t-statistic
    n1 <- sample1$Num_Replicates
    n2 <- sample2$Num_Replicates
    s1 <- sample1$PropagatedError
    s2 <- sample2$PropagatedError
    x1 <- sample1$mean
    x2 <- sample2$mean
    
    t_statistic <- (x1 - x2) / sqrt(s1^2 / n1 + s2^2 / n2)
    
    # Calculate degrees of freedom
    df <- ((s1^2 / n1 + s2^2 / n2)^2) /
      ((s1^2 / n1)^2 / (n1 - 1) + (s2^2 / n2)^2 / (n2 - 1))
    
    # Determine the p-value
    p_value <- 2 * pt(-abs(t_statistic), df)
    
    return(list(samples = paste(combo, collapse = " vs "), 
                t_statistic = t_statistic, p.value = p_value, df = df))
  })
  return(do.call(rbind, results))
}

# Perform the tests for both IPTG concentrations
results_0 <- pairwise_t_test(data_0)
results_1000 <- pairwise_t_test(data_1000)


# Function to perform paired t-test
paired_t_test <- function(data) {
  # Extract unique samples
  unique_samples <- unique(data$Sample)
  
  results <- data.frame(Sample = character(), mean_difference = numeric(), 
                        t_statistic = numeric(), 
                        p_value = numeric(), 
                        df = numeric(), stringsAsFactors = FALSE)
  
  for (sample in unique_samples) {
    # Filter data for this sample at both IPTG concentrations
    sample_data <- subset(data, Sample == sample)
    
    if (nrow(sample_data) == 2) { # Ensure we have both IPTG = 0 and IPTG = 1000
      # Extract data for IPTG = 0 and IPTG = 1000
      d0 <- subset(sample_data, IPTG_conc == 0)
      d1000 <- subset(sample_data, IPTG_conc == 1000)
      
      # Calculate mean difference
      mean_diff <- d0$mean - d1000$mean
      
      # Calculate the combined standard error of the difference
      se_diff <- sqrt(d0$PropagatedError^2 + d1000$PropagatedError^2)
      
      # Calculate t-statistic
      t_statistic <- mean_diff / se_diff
      
      # Number of pairs is given by the number of replicates 
      #(minimum from both groups to ensure matching pairs)
      n <- min(d0$Num_Replicates, d1000$Num_Replicates)
      
      # Degrees of freedom: n - 1
      df <- n - 1
      
      # Calculate p-value
      if (df > 0) {
        p_value <- 2 * pt(-abs(t_statistic), df)
      } else {
        p_value <- NA  # Not defined for df <= 0
      }
      
      # Append results
      results <- rbind(results, data.frame(Sample = sample, 
                                           mean_difference = mean_diff, 
                                           t_statistic = t_statistic, 
                                           p_value = p_value, df = df))
    }
  }
  
  return(results)
}

# Perform the paired t-test
paired_results <- paired_t_test(error_propagation)
```

#### Final stats tests with p-values

Within promoter (0 vs 1 mM IPTG), paired t-test results:

```{r echo=FALSE}
# Print the results
print(paired_results)
```

Welch's t-tests for promoters with no induction (0mM IPTG):

```{r echo=FALSE}
print(results_0)
```

Welch's t-tests for promoters at full induction (1mM IPTG):

```{r echo=FALSE}
print(results_1000)
```

## Titration of sfGFP expression across IPTG concentrations (Fig 3/S2)

#### Plasmid expression

Load cleaned data from plate reader

```{r echo=FALSE}
ind_exp_data <- read.csv("pAb35_sfGFP_IPTG_curve.csv", header = TRUE)

print(ind_exp_data)
```

Subtract background values and normalize to cell density (fluorescence/OD)

```{r}
columns_to_normalize <- c('X19606.fluo', 'X19606.OD', 'X5075.fluo', 'X5075.OD', 
                          'X17978.fluo', 'X17978.OD', 'Eco.fluo', 'Eco.OD')

background_vals <- ind_exp_data %>% filter(IPTG_conc %like% "background") %>% 
  select(all_of(columns_to_normalize))

# Subtracting background averages from each column
for (col in columns_to_normalize) {
  ind_exp_data[[col]] <- ind_exp_data[[col]] - background_vals[[col]]
}

# Remove the background_average row if no longer needed
ind_exp_data <- ind_exp_data %>% filter(!(IPTG_conc %like% "background"))

# Identifying pairs of columns for division
fluo_columns <- grep("\\.fluo$", names(ind_exp_data), value = TRUE)
od_columns <- sub("fluo", "OD", fluo_columns)

# Performing division and storing results in new columns
for (i in seq_along(fluo_columns)) {
  new_col_name <- paste0(sub("\\.fluo", "", fluo_columns[i]), ".Ratio")
  ind_exp_data[[new_col_name]] <- ind_exp_data[[fluo_columns[i]]] / 
    ind_exp_data[[od_columns[i]]]
}

data.norm <- ind_exp_data %>%
  select(IPTG_conc, rep, ends_with(".Ratio"))
```

Subtract autofluorescence (empty vector)\
Additionally, propagate error for standard deviation

```{r}
# Filter data for IPTG_conc = "EV" and calculate mean and SD
ev_stats <- data.norm %>%
  filter(IPTG_conc == "EV") %>%
  summarise(across(ends_with(".Ratio"), list(mean = mean, sd = sd)))

# Calculate mean and SD of the samples
mean_sd_diff <- data.norm %>%
  group_by(IPTG_conc) %>%
  summarise(across(ends_with(".Ratio"), list(mean = mean, sd = sd))) %>%
  filter(IPTG_conc != "EV")

# Subtract mean EV from other values in each .Ratio column
data_adjusted <- data.norm %>%
  filter(IPTG_conc != "EV") %>%
  rowwise() %>%
  mutate(
    X19606.Ratio = X19606.Ratio - ev_stats$X19606.Ratio_mean,
    X5075.Ratio = X5075.Ratio - ev_stats$X5075.Ratio_mean,
    X17978.Ratio = X17978.Ratio - ev_stats$X17978.Ratio_mean,
    Eco.Ratio = Eco.Ratio - ev_stats$Eco.Ratio_mean
  )

mean_sd_adjusted <- data_adjusted %>%
  group_by(IPTG_conc) %>%
  summarise(across(ends_with(".Ratio"), list(mean = mean, sd = sd)))

# Propagate error (assuming errors are uncorrelated)
error_propagation <- mean_sd_adjusted %>%
  mutate(
    X19606.Ratio_PropagatedError = sqrt(mean_sd_diff$X19606.Ratio_sd^2 + 
                                          ev_stats$X19606.Ratio_sd^2),
    X5075.Ratio_PropagatedError = sqrt(mean_sd_diff$X5075.Ratio_sd^2 + 
                                         ev_stats$X5075.Ratio_sd^2),
    X17978.Ratio_PropagatedError = sqrt(mean_sd_diff$X17978.Ratio_sd^2 + 
                                          ev_stats$X17978.Ratio_sd^2),
    Eco.Ratio_PropagatedError = sqrt(mean_sd_diff$Eco.Ratio_sd^2 + 
                                       ev_stats$Eco.Ratio_sd^2)
  ) %>%
  select(IPTG_conc, ends_with("_mean"), ends_with("PropagatedError")) %>%
  pivot_longer(
    cols = -IPTG_conc, 
    names_to = c("Sample", ".value"),
    names_pattern = "(.*)(mean|PropagatedError)$"
  ) %>%
  mutate(Sample = str_remove(Sample, "_"))
```

```{r echo=FALSE}
print(error_propagation)
```

Plot expression with IPTG concentration on a semilog scale

##### *A. baumannii* ATCC 17978 and *E. coli* BW25113

```{r}
# Select for strains and determine appropriate limits for the axes
mean_sd_data <- error_propagation %>% 
  filter(Sample %like% "17978" | Sample %like% "Eco") 

mean_sd_data$IPTG_conc <- as.numeric(mean_sd_data$IPTG_conc)

x_limits <- range(mean_sd_data$IPTG_conc, na.rm = TRUE)
y_limits <- mean_sd_data %>%
  mutate(
    Lower = pmin(0, mean - PropagatedError),
    Upper = mean + PropagatedError
  ) %>%
  summarise(
    Min = min(Lower, na.rm = TRUE),
    Max = max(Upper, na.rm = TRUE)
  ) %>%
  unlist() 
```

```{r}
ggplot(mean_sd_data, aes(x = IPTG_conc, y = mean, group = Sample)) +
  geom_point(aes(color = Sample)) +
  geom_line(aes(color = Sample)) +
  geom_errorbar(aes(ymin = mean - PropagatedError, 
                    ymax = mean + PropagatedError, color = Sample), 
                width = 0.02) +
  geom_vline(xintercept = 0, color = "black", size = 1) + 
  labs(x = "IPTG Concentration (uM)", 
       y = "Expression (Normalized Fluorescence)", 
       title = "GFP plasmid vector") +
  theme_minimal() +
  scale_x_log10() +
  scale_y_continuous(limits = y_limits)  +
  scale_color_manual(values = c("Eco.Ratio" = "lightgreen", "X17978.Ratio" = "red3"))+
  theme(legend.position = "bottom")
```

##### *A. baumannii* ATCC 19606 and AB5075

```{r}
# Select for strains and determine appropriate limits for the axes
mean_sd_data <- error_propagation %>% 
  filter(Sample %like% "5075" | Sample %like% "19606") 

mean_sd_data$IPTG_conc <- as.numeric(mean_sd_data$IPTG_conc)

x_limits <- range(mean_sd_data$IPTG_conc, na.rm = TRUE)
y_limits <- mean_sd_data %>%
  mutate(
    Lower = pmin(0, mean - PropagatedError),
    Upper = mean + PropagatedError
  ) %>%
  summarise(
    Min = min(Lower, na.rm = TRUE),
    Max = max(Upper, na.rm = TRUE)
  ) %>%
  unlist() 
```

```{r}
ggplot(mean_sd_data, aes(x = IPTG_conc, y = mean, group = Sample)) +
  geom_point(aes(color = Sample)) +
  geom_line(aes(color = Sample)) +
  geom_errorbar(aes(ymin = mean - PropagatedError, 
                    ymax = mean + PropagatedError, 
                    color = Sample), width = 0.02) +
  geom_vline(xintercept = 0, color = "black", size = 1) +  
  labs(x = "IPTG Concentration (uM)", 
       y = "Expression (Normalized Fluorescence)", 
       title = "GFP Plasmid Induction-Expression") +
  theme_minimal() +
  scale_x_log10() +
  scale_y_continuous(limits = y_limits)  +
  scale_color_manual(values = c("X19606.Ratio" = "plum1", "X5075.Ratio" = "purple3"))+
  theme(legend.position = "bottom")
```

#### Tn7 transposon Expression

Load cleaned data from plate reader

```{r echo=FALSE}
ind_exp_Tn7data <- fread("./Tn7/240424_Tn7_expression_100.csv", header = TRUE)

print(ind_exp_Tn7data)
```

Subtract background values and normalize to cell density (fluorescence/OD)

```{r}
columns_to_normalize <- c('19606.fluo', '19606.OD', '17978.fluo', '17978.OD', 
                          'Eco.fluo', 'Eco.OD','5075.fluo', '5075.OD')

# Getting the background average values
background_vals <- ind_exp_Tn7data %>% filter(IPTG_conc %like% "background") %>% 
  select(all_of(columns_to_normalize))

# Subtracting background averages from each column
for (col in columns_to_normalize) {
  ind_exp_Tn7data[[col]] <- ind_exp_Tn7data[[col]] - background_vals[[col]]
}

# Remove the background_average row if no longer needed
ind_exp_Tn7data <- ind_exp_Tn7data %>% filter(!(IPTG_conc %like% "background"))

# Identifying pairs of columns for division
fluo_columns <- grep("\\.fluo$", names(ind_exp_Tn7data), value = TRUE)
od_columns <- sub("fluo", "OD", fluo_columns)

# Performing division and storing results in new columns
for (i in seq_along(fluo_columns)) {
  new_col_name <- paste0(sub("\\.fluo", "", fluo_columns[i]), ".Ratio")
  ind_exp_Tn7data[[new_col_name]] <- ind_exp_Tn7data[[fluo_columns[i]]] / 
    ind_exp_Tn7data[[od_columns[i]]]
}

data.norm <- ind_exp_Tn7data %>%
  select(IPTG_conc, rep, ends_with(".Ratio"))
```

Subtract autofluorescence (empty vector)
Additionally, propagate error for standard deviation

```{r}
# Filter data for IPTG_conc = "EV" and calculate mean and SD
ev_stats <- data.norm %>%
  filter(IPTG_conc == "EV") %>%
  summarise(across(ends_with(".Ratio"), list(mean = mean, sd = sd)))

# Calculate mean and SD of the sample data
mean_sd_diff <- data.norm %>%
  group_by(IPTG_conc) %>%
  summarise(across(ends_with(".Ratio"), list(mean = mean, sd = sd))) %>%
  filter(IPTG_conc != "EV")

# Subtract mean EV from other values in each .Ratio column
data_adjusted <- data.norm %>%
  filter(IPTG_conc != "EV") %>%
  rowwise() %>%
  mutate(
    `19606.Ratio` = `19606.Ratio` - ev_stats$`19606.Ratio_mean`,
    `17978.Ratio` = `17978.Ratio` - ev_stats$`17978.Ratio_mean`,
    Eco.Ratio = Eco.Ratio - ev_stats$Eco.Ratio_mean,
    `5075.Ratio` = `5075.Ratio` - ev_stats$`5075.Ratio_mean`
  )

mean_sd_adjusted <- data_adjusted %>%
  group_by(IPTG_conc) %>%
  summarise(across(ends_with(".Ratio"), list(mean = mean, sd = sd)))

# Propagate error (assuming errors are uncorrelated)
error_propagation <- mean_sd_adjusted %>%
  mutate(
    `19606.Ratio_PropagatedError` = sqrt(mean_sd_diff$`19606.Ratio_sd`^2 + 
                                           ev_stats$`19606.Ratio_sd`^2),
    `17978.Ratio_PropagatedError` = sqrt(mean_sd_diff$`17978.Ratio_sd`^2 + 
                                           ev_stats$`17978.Ratio_sd`^2),
    Eco.Ratio_PropagatedError = sqrt(mean_sd_diff$Eco.Ratio_sd^2 + 
                                       ev_stats$Eco.Ratio_sd^2),
    `5075.Ratio_PropagatedError` = sqrt(mean_sd_diff$`5075.Ratio_sd`^2 + 
                                          ev_stats$`5075.Ratio_sd`^2)) %>%
  select(IPTG_conc, ends_with("_mean"), ends_with("PropagatedError")) %>%
  pivot_longer(
    cols = -IPTG_conc, 
    names_to = c("Sample", ".value"),
    names_pattern = "(.*)(mean|PropagatedError)$"
  ) %>%
  mutate(Sample = str_remove(Sample, "_"))
```

```{r echo=FALSE}
print(error_propagation)
```

Plot expression levels across inducer (on semilog axis)

##### *A. baumannii* ATCC 17978 and *E. coli* BW25113

```{r}
# Select for strains and determine appropriate limits for the axes
mean_sd_data <- error_propagation %>% 
  filter(Sample %like% "17978" | Sample %like% "Eco") 

mean_sd_data$IPTG_conc <- as.numeric(mean_sd_data$IPTG_conc)

x_limits <- range(mean_sd_data$IPTG_conc, na.rm = TRUE)
y_limits <- mean_sd_data %>%
  mutate(
    Lower = pmin(0, mean - PropagatedError),
    Upper = mean + PropagatedError
  ) %>%
  summarise(
    Min = min(Lower, na.rm = TRUE),
    Max = max(Upper, na.rm = TRUE)
  ) %>%
  unlist() 
```

```{r}
ggplot(mean_sd_data, aes(x = IPTG_conc, y = mean, group = Sample)) +
  geom_point(aes(color = Sample)) +
  geom_line(aes(color = Sample)) +
  geom_errorbar(aes(ymin = mean - PropagatedError, 
                    ymax = mean + PropagatedError, 
                    color = Sample), width = 0.02) +
  geom_vline(xintercept = 0, color = "black", linewidth = 1) +  
  labs(x = "IPTG Concentration (uM)", 
       y = "Expression (Normalized Fluorescence)", 
       title = "Tn7 vector Induction-Expression") +
  theme_minimal() +
  scale_x_log10() +
  scale_y_continuous(limits = y_limits)  +
  scale_color_manual(values = c("Eco.Ratio" = "lightgreen", 
                                "17978.Ratio" = "red3"))+
  theme(legend.position = "bottom")
```

##### *A. baumannii* ATCC 19606 and AB5075

```{r}
# Select for strains and determine appropriate limits for the axes
mean_sd_data <- error_propagation %>% 
  filter(Sample %like% "19606" | Sample %like% "5075") 

mean_sd_data$IPTG_conc <- as.numeric(mean_sd_data$IPTG_conc)

x_limits <- range(mean_sd_data$IPTG_conc, na.rm = TRUE)
y_limits <- mean_sd_data %>%
  mutate(
    Lower = pmin(0, mean - PropagatedError),
    Upper = mean + PropagatedError
  ) %>%
  summarise(
    Min = min(Lower, na.rm = TRUE),
    Max = max(Upper, na.rm = TRUE)
  ) %>%
  unlist() 
```

```{r}
ggplot(mean_sd_data, aes(x = IPTG_conc, y = mean, group = Sample)) +
  geom_point(aes(color = Sample)) +
  geom_line(aes(color = Sample)) +
  geom_errorbar(aes(ymin = mean - PropagatedError, 
                    ymax = mean + PropagatedError, 
                    color = Sample), width = 0.02) +
  geom_vline(xintercept = 0, color = "black", linewidth = 1) +
  labs(x = "IPTG Concentration (uM)", 
       y = "Expression (Normalized Fluorescence)", 
       title = "Tn7 vector Induction-Expression") +
  theme_minimal() +
  scale_x_log10() +
  scale_y_continuous(limits = y_limits)  +
  scale_color_manual(values = c("19606.Ratio" = "plum1", 
                                "5075.Ratio" = "purple3"))+
  theme(legend.position = "bottom")
```


