---
title: "rpoE graphs"
author: "Jennifer Tran"
date: "2024-08-19"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## RpoE fluorescence assay (Fig 5)

Load necessary packages for these graphs:

```{r, warning=FALSE, message = FALSE}
require('pacman')

p_load(dplyr, ggplot2, tidyr, RColorBrewer)
```

Load reformatted plate data 

```{r, echo=FALSE}
rpoE_data <- read.csv("./20240521_RpoE_data.csv", header = TRUE)

print(rpoE_data)
```

Subtract off background fluorescence and normalize to OD
```{r}
background_vals <- rpoE_data %>% filter(IPTG_conc == "background")

rpoE_norm <- rpoE_data %>% filter(IPTG_conc != "background") %>% 
  mutate(IPTG_conc = as.numeric(IPTG_conc),
         fluorescence = fluorescence - background_vals$fluorescence,
         OD = OD - background_vals$OD)

rpoE_norm$norm_exp <- rpoE_norm$fluorescence/rpoE_norm$OD
```

Center autofluorescence (WT, no expression vectors) around 0 and get stats
```{r, message = FALSE}
WT_stats <- rpoE_norm %>% filter(IPTG_conc == 0) %>%
  group_by(strain) %>%
  summarise(WTmean = mean(norm_exp, na.rm = TRUE),
            WTsd = sd(norm_exp, na.rm = TRUE),
            N = n_distinct(rep))

sample_stats <- rpoE_norm %>% 
  group_by(IPTG_conc, strain, plasmid) %>%
  summarise(mean = mean(norm_exp, na.rm = TRUE),
            sd = sd(norm_exp, na.rm = TRUE),
            N = n_distinct(rep))

rpoE_adjusted <- rpoE_norm %>%
  left_join(WT_stats, by = "strain") %>%
  mutate(combination = interaction(IPTG_conc, plasmid),
    norm_exp_adj = norm_exp - WTmean) %>%
  select(-WTmean, -WTsd)

combined_stats <- sample_stats %>%
  left_join(WT_stats, by = "strain") %>%
  mutate(sd_adj = sqrt(WTsd^2 + sd^2),
         mean_adj = mean - WTmean,
         N_final = ifelse(N.x == N.y, N.x, NA))

```

Plot bar graphs with error-propagated standard deviations

#####for _A. baumannii_
```{r}
ab_points <- rpoE_adjusted %>%
  filter(strain == "Ab") %>% filter(IPTG_conc != 0.050)

combined_stats %>% filter(strain == "Ab") %>% filter(IPTG_conc != 0.05) %>%
  mutate(combination = interaction(IPTG_conc, plasmid)) %>%
  ggplot(aes(x = combination, y = mean_adj)) +
  geom_bar(stat = "identity", fill = "pink3", color = "black") +
  geom_errorbar(aes(ymin = mean_adj - sd_adj, 
                    ymax = mean_adj + sd_adj), width = 0.5) +
  geom_point(data = ab_points, aes(y=norm_exp_adj), 
             position = position_jitter(width = 0.2), size = 2) +
  ylim(limits = c(-210, 60000)) +
  labs(x = "Sample", y = "mRFP expression (normalized fluorescence)",
       title = "A. baumannii PrpoE reporter assay") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#####and for _E. coli_
```{r}
eco_points <- rpoE_adjusted %>%
  filter(strain == "Eco")

combined_stats %>% filter(strain == "Eco") %>%
  mutate(combination = interaction(IPTG_conc, plasmid)) %>%
  ggplot(aes(x = combination, y = mean_adj)) +
  geom_bar(stat = "identity", fill = "lightgreen", color = "black") +
  geom_errorbar(aes(ymin = mean_adj - sd_adj, 
                    ymax = mean_adj + sd_adj), width = 0.5) +
  geom_point(data = eco_points, aes(y=norm_exp_adj), 
             position = position_jitter(width = 0.2), size = 2) +
  labs(x = "Sample", y = "mRFP expression (normalized fluorescence)",
       title = "E. coli PrpoE reporter assay") +
  ylim(limits = c(-210, 60000)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


And run statistics to determine differences (Welch's t-tests)

```{r}
# Function to perform pairwise unpaired t-tests
perform_pairwise_t_tests <- function(data) {
  plasmids <- unique(data$plasmid)
  results <- list()
  num_comparisons <- length(plasmids) * (length(plasmids) - 1) / 2

  for (i in 1:(length(plasmids) - 1)) {
    for (j in (i + 1):length(plasmids)) {
      plasmid1 <- plasmids[i]
      plasmid2 <- plasmids[j]
      
      data1 <- data %>% filter(plasmid == plasmid1)
      data2 <- data %>% filter(plasmid == plasmid2)

      # Ensure that the number of rows in each group being compared is the same
      if(nrow(data1) == nrow(data2)) {
        # Extract values
        mean1 <- data1$mean_adj
        mean2 <- data2$mean_adj
        sd1 <- data1$sd_adj
        sd2 <- data2$sd_adj
        n1 <- data1$N_final
        n2 <- data2$N_final
        
        # Calculate standard error of the difference
        se_diff <- sqrt(sd1^2 / n1 + sd2^2 / n2)
        
        if(all(se_diff > 0)) { # Ensure that no division by zero occurs
          # Calculate Welch's t-statistic
          t_statistic <- (mean1 - mean2) / se_diff

          # Calculate degrees of freedom for Welch's t-test
          num <- (sd1^2 / n1 + sd2^2 / n2)^2
          denom <- ((sd1^2 / n1)^2 / (n1 - 1)) + ((sd2^2 / n2)^2 / (n2 - 1))
          df <- num / denom
          
          p_value <- 2 * pt(-abs(t_statistic), df)
          bonferroni_adj <- min(p_value * num_comparisons, 1)  # Bonferroni adjustment
          
          # Store results
          results[[paste(plasmid1, plasmid2, sep = "_vs_")]] <- data.frame(
            Plasmid1 = plasmid1,
            Plasmid2 = plasmid2,
            t_statistic = t_statistic,
            SEM = se_diff,
            degrees_of_freedom = df,
            p_value = p_value,
            Bonferroni_adj = bonferroni_adj
          )
        }
      }
    }
  }

  do.call(rbind, results)
}

# Filter the data for Ab and Eco strains
ab_data <- combined_stats %>% filter(strain == "Ab") %>% 
  filter(IPTG_conc != 0.050) 

eco_data <- filter(combined_stats, strain == "Eco") 

# Perform pairwise comparisons within each strain
ab_comparisons <- perform_pairwise_t_tests(ab_data)
eco_comparisons <- perform_pairwise_t_tests(eco_data)
```

##### Welch's t-test results
```{r, echo=FALSE}
print("_A. baumannii_ stats")
print(ab_comparisons)
```
```{r, echo=FALSE}
print("_E. coli_ stats")
print(eco_comparisons)
```


## Other RpoE-regulated promoters fluorescence assay (Fig S4)


Load reformatted plate data 

```{r, echo=FALSE}
prom_data <- read.csv("./rpoE/20240815_rpoE_data.csv", header = TRUE)

print(prom_data)
```

Subtract off background fluorescence and OD and normalize
```{r}
background_prom_vals <- prom_data %>% filter(strain == "background")

prom_norm <- prom_data %>% filter(strain != "background") %>% 
  mutate(fluorescence = fluorescence - background_prom_vals$fluorescence,
         OD = OD - background_prom_vals$OD)

prom_norm$norm_exp <- prom_norm$fluorescence/prom_norm$OD
```

Center background fluorescence (no promoter ahead of mRFP) around 0 and get stats
```{r, message = FALSE}
Pempty_stats <- prom_norm %>% filter(Tn7 == "Pempty") %>%
  reframe(strain = 'Ab',
            Pemptymean = mean(norm_exp, na.rm = TRUE),
            Pemptysd = sd(norm_exp, na.rm = TRUE),
            N = n_distinct(rep))

sample_prom_stats <- prom_norm %>% 
  group_by(plasmid, Tn7) %>%
  summarise(strain = "Ab",
            mean = mean(norm_exp, na.rm = TRUE),
            sd = sd(norm_exp, na.rm = TRUE),
            N = n_distinct(rep))

prom_adjusted <- prom_norm %>%
  left_join(Pempty_stats, by = "strain") %>%
  mutate(combination = interaction(plasmid, Tn7),
    norm_exp_adj = norm_exp - Pemptymean) %>%
  select(-Pemptymean, -Pemptysd)

combined_prom_stats <- sample_prom_stats %>%
  left_join(Pempty_stats, by = "strain") %>%
  mutate(sd_adj = sqrt(Pemptysd^2 + sd^2),
         mean_adj = mean - Pemptymean,
         N_final = ifelse(N.x == N.y, N.x, NA))

```

Plot bar graphs with error-propagated standard deviations


```{r}
ggplot(combined_prom_stats, aes(x = Tn7, y = mean_adj, fill = plasmid)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  geom_errorbar(aes(ymin = mean_adj - sd_adj, ymax = mean_adj + sd_adj),
                width = 0.5, position = position_dodge(0.9)) +
  geom_point(data = prom_adjusted, aes(x = Tn7, y = norm_exp_adj, group = plasmid), 
             position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.9), size = 2, color = "black") +
  labs(x = "Promoter", y = "mRFP expression (normalized fluorescence)", title = "RpoE-dependent promoter expression") +
  scale_y_continuous(limits = c(-33, 75000)) +
  scale_fill_manual(values = c("RpoE" = "salmon", "EV" = "grey", "none" = "gray30")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")
```


And run statistics to determine differences (Welch's t-tests)

```{r}

# Filter the data for plasmids
PmicA_data <- filter(combined_prom_stats, Tn7 == "PmicA" | Tn7 == "Pempty")  

PrybB_data <- filter(combined_prom_stats, Tn7 == "PrybB"| Tn7 == "Pempty") 

PyicJ_data <- filter(combined_prom_stats, Tn7 == "PyicJ"| Tn7 == "Pempty") 


# Perform pairwise comparisons within each strain using previously defined function
PmicA_comparisons <- perform_pairwise_t_tests(PmicA_data)
PrybB_comparisons <- perform_pairwise_t_tests(PrybB_data)
PyicJ_comparisons <- perform_pairwise_t_tests(PyicJ_data)


```

##### Welch's t-test results
```{r, echo=FALSE}
print("PmicA stats")
print(PmicA_comparisons)
```
```{r, echo=FALSE}
print("PrybB stats")
print(PrybB_comparisons)
```
```{r, echo=FALSE}
print("PyicJ stats")
print(PyicJ_comparisons)
```