---
title: "Single cell expression analyses"
author: "Jennifer Tran"
date: "2024-08-19"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load necessary packages for these graphs:

```{r, warning=FALSE, message = FALSE}
require('pacman')

p_load(dplyr, data.table, ggplot2, tidyr, RColorBrewer, stringr, ggridges, colourpicker, ggforce)
```

## Flow data ridgeplots for plasmid expression (Fig 4)

Load all data tables (.csv for each sample) and sample names

```{r echo=FALSE}
aba01 <- fread("C:/Users/jentr/OneDrive - UW-Madison/lab things/data/Ab_plasmids/export_Specimen_001_aba_001.csv", header = TRUE)
aba02 <- fread("C:/Users/jentr/OneDrive - UW-Madison/lab things/data/Ab_plasmids/export_Specimen_001_aba_002.csv", header = TRUE)
aba03 <- fread("C:/Users/jentr/OneDrive - UW-Madison/lab things/data/Ab_plasmids/export_Specimen_001_aba_003.csv", header = TRUE)
aba04 <- fread("C:/Users/jentr/OneDrive - UW-Madison/lab things/data/Ab_plasmids/export_Specimen_001_aba_004.csv", header = TRUE)
aba05 <- fread("C:/Users/jentr/OneDrive - UW-Madison/lab things/data/Ab_plasmids/export_Specimen_001_aba_005.csv", header = TRUE)
aba06 <- fread("C:/Users/jentr/OneDrive - UW-Madison/lab things/data/Ab_plasmids/export_Specimen_001_aba_006.csv", header = TRUE)
aba07 <- fread("C:/Users/jentr/OneDrive - UW-Madison/lab things/data/Ab_plasmids/export_Specimen_001_aba_007.csv", header = TRUE)
aba08 <- fread("C:/Users/jentr/OneDrive - UW-Madison/lab things/data/Ab_plasmids/export_Specimen_001_aba_008.csv", header = TRUE)
aba09 <- fread("C:/Users/jentr/OneDrive - UW-Madison/lab things/data/Ab_plasmids/export_Specimen_001_aba_009.csv", header = TRUE)
aba10 <- fread("C:/Users/jentr/OneDrive - UW-Madison/lab things/data/Ab_plasmids/export_Specimen_001_aba_010.csv", header = TRUE)

eco01 <- fread("C:/Users/jentr/OneDrive - UW-Madison/lab things/data/Ab_plasmids/export_Specimen_001_eco_1.csv", header = TRUE)
eco02 <- fread("C:/Users/jentr/OneDrive - UW-Madison/lab things/data/Ab_plasmids/export_Specimen_001_eco_2.csv", header = TRUE)
eco03 <- fread("C:/Users/jentr/OneDrive - UW-Madison/lab things/data/Ab_plasmids/export_Specimen_001_eco_3.csv", header = TRUE)
eco04 <- fread("C:/Users/jentr/OneDrive - UW-Madison/lab things/data/Ab_plasmids/export_Specimen_001_eco_4.csv", header = TRUE)
eco05 <- fread("C:/Users/jentr/OneDrive - UW-Madison/lab things/data/Ab_plasmids/export_Specimen_001_eco_5.csv", header = TRUE)
eco06 <- fread("C:/Users/jentr/OneDrive - UW-Madison/lab things/data/Ab_plasmids/export_Specimen_001_eco_6.csv", header = TRUE)
eco07 <- fread("C:/Users/jentr/OneDrive - UW-Madison/lab things/data/Ab_plasmids/export_Specimen_001_eco_7.csv", header = TRUE)
eco08 <- fread("C:/Users/jentr/OneDrive - UW-Madison/lab things/data/Ab_plasmids/export_Specimen_001_eco_8.csv", header = TRUE)
eco09 <- fread("C:/Users/jentr/OneDrive - UW-Madison/lab things/data/Ab_plasmids/export_Specimen_001_eco_9.csv", header = TRUE)
eco10 <- fread("C:/Users/jentr/OneDrive - UW-Madison/lab things/data/Ab_plasmids/export_Specimen_001_eco_10.csv", header = TRUE)

sample_map <- fread("C:/Users/jentr/OneDrive - UW-Madison/lab things/data/Ab_plasmids/20240405_sample_ID.csv", header = TRUE)

print(sample_map)
```

Place all loaded dataframes into a list

```{r}
#A. baumannii list
list_of_aba <- list(aba01 = aba01, aba02 = aba02, aba03 = aba03, aba04 = aba04, 
                    aba05 = aba05, aba06 = aba06, aba07 = aba07, aba08 = aba08, 
                    aba09 = aba09, aba10 = aba10)

combined_aba <- rbindlist(list_of_aba, idcol = "SampleID")

#E. coli list
list_of_eco <- list(eco01 = eco01, eco02 = eco02, eco03 = eco03, eco04 = eco04, 
                    eco05 = eco05, eco06 = eco06, eco07 = eco07, eco08 = eco08,
                    eco09 = eco09, eco10 = eco10)

combined_eco <- rbindlist(list_of_eco, idcol = "SampleID")
```

```{r echo=FALSE}
print("A. baumannii list")
glimpse(combined_aba)
print("E. coli list")
glimpse(combined_eco)

```

Normalize fluorescence data (GFP readings) to empty vector control\
This centers no fluorescence readings around 0; relative fluorescence


```{r}
##for A. baumannii
# Filter the negative control for A. baumannii
neg_control_ab <- combined_aba %>% filter(SampleID == "aba01") 

# Calculate the mean and standard deviation for the negative control
mean_ab <- mean(neg_control_ab$GFP, na.rm = TRUE)
sd_ab <- sd(neg_control_ab$GFP, na.rm = TRUE)

# Normalize GFP values by subtracting the mean of the negative control
combined_aba <- combined_aba %>%
  mutate(norm_GFP = GFP - mean_ab)

# Add IDs from the sample map
combined_aba <- combined_aba %>% 
  left_join(sample_map, by = "SampleID")

# Create a factor for IPTG concentration levels
combined_aba$IPTG_uM_factor <- factor(
  combined_aba$IPTG_uM, 
  levels = unique(combined_aba$IPTG_uM[order(combined_aba$SampleID)])
)

# Add IDs from the sample map and calculate statistics for each sample
combined_aba_stats <- combined_aba %>% 
  group_by(SampleID) %>%
  summarise(
    mean = mean(GFP, na.rm = TRUE),
    sd = sd(GFP, na.rm = TRUE),
    N = n(),
    .groups = 'drop'
  ) %>%
  mutate(
    sd_adj = sqrt(sd_ab^2 + sd^2),
    mean_adj = mean - mean_ab,
    N_final = N
  ) %>%
  left_join(sample_map, by = "SampleID")

print(combined_aba_stats)
```
```{r}
##for E. coli##
# Filter the negative control for E. coli
neg_control_eco <- combined_eco %>% filter(SampleID == "eco01") 

# Calculate the mean and standard deviation for the negative control
mean_eco <- mean(neg_control_eco$GFP, na.rm = TRUE)
sd_eco <- sd(neg_control_eco$GFP, na.rm = TRUE)
neg_n_eco <- nrow(neg_control_eco)

# Normalize GFP values by subtracting the mean of the negative control
combined_eco <- combined_eco %>%
  mutate(norm_GFP = GFP - mean_eco)

# Add IDs from the sample map
combined_eco <- combined_eco %>% 
  left_join(sample_map, by = "SampleID")

# Create a factor for IPTG concentration levels
combined_eco$IPTG_uM_factor <- factor(
  combined_eco$IPTG_uM, 
  levels = unique(combined_eco$IPTG_uM[order(combined_eco$SampleID)])
)

# Add IDs from the sample map and calculate statistics for each sample
combined_eco_stats <- combined_eco %>% 
  group_by(SampleID) %>%
  summarise(
    mean = mean(GFP, na.rm = TRUE),
    sd = sd(GFP, na.rm = TRUE),
    N = n(),
    .groups = 'drop'
  ) %>%
  mutate(
    sd_adj = sqrt(sd_eco^2 + sd^2),
    mean_adj = mean - mean_eco,
    N_final = N
  ) %>%
    left_join(sample_map, by = "SampleID")

print(combined_eco_stats)


```

Plot stacked density plots for each IPTG concentration

#####*A. baumannii* ATCC 17978

```{r}
ggplot(combined_aba, aes(x = norm_GFP, y = IPTG_uM_factor, 
                         height = after_stat(density))) +
  geom_density_ridges_gradient(
    scale = 2, # Could adjust for scale
    aes(fill = after_stat(x)), 
    gradient_lwd = 0.0, 
    rel_min_height = 0.02 
  ) +
  scale_fill_viridis_c(name = "norm_GFP", option = "G") +
  labs(x = "GFP Expression (Fluorescence)", y = "IPTG concentration (uM)", 
       title = "GFP Plasmid - A. baumannii") +
  theme_ridges() + theme(legend.position = "none") 
```

#####for *E. coli* BW25113

```{r}
ggplot(combined_eco, aes(x = norm_GFP, y = IPTG_uM_factor, height = after_stat(density))) +
  geom_density_ridges_gradient(
    scale = 2, 
    aes(fill = after_stat(x)), 
    gradient_lwd = 0.0, 
    rel_min_height = 0.02 
  ) +
  scale_fill_viridis_c(name = "norm_GFP", option = "D") +
  labs(x = "GFP Expression (Fluorescence)", y = "IPTG concentration (uM)", 
       title = "GFP Plasmid - E. coli") +
  theme_ridges() + theme(legend.position = "none")
```



```{r}
# Function to perform pairwise unpaired t-tests with Bonferroni adjustment
perform_pairwise_t_tests <- function(data) {
  IPTG_conc <- unique(data$IPTG_uM)
  results <- list()
  num_comparisons <- length(IPTG_conc) * (length(IPTG_conc) - 1) / 2  # Total number of comparisons

  for (i in 1:(length(IPTG_conc) - 1)) {
    for (j in (i + 1):length(IPTG_conc)) {
      IPTG_conc1 <- IPTG_conc[i]
      IPTG_conc2 <- IPTG_conc[j]
      
      data1 <- data %>% filter(IPTG_uM == IPTG_conc1)
      data2 <- data %>% filter(IPTG_uM == IPTG_conc2)

      # Ensure that the number of rows in each group being compared is the same
      if(nrow(data1) == nrow(data2)) {
        # Extract values
        mean1 <- data1$mean_adj
        mean2 <- data2$mean_adj
        sd1 <- data1$sd_adj
        sd2 <- data2$sd_adj
        n1 <- data1$N_final
        n2 <- data2$N_final
        
        # Calculate standard error of the difference
        se_diff <- sqrt(sd1^2 / n1 + sd2^2 / n2)
        
        if(all(se_diff > 0)) { # Ensure that no division by zero occurs
          # Calculate Welch's t-statistic
          t_statistic <- (mean1 - mean2) / se_diff

          # Calculate degrees of freedom for Welch's t-test
          num <- (sd1^2 / n1 + sd2^2 / n2)^2
          denom <- ((sd1^2 / n1)^2 / (n1 - 1)) + ((sd2^2 / n2)^2 / (n2 - 1))
          df <- num / denom
          
          p_value <- 2 * pt(-abs(t_statistic), df)
          bonferroni_adj <- min(p_value * num_comparisons, 1)  # Bonferroni adjustment
          
          # Store results
          results[[paste(IPTG_conc1, IPTG_conc2, sep = "_vs_")]] <- data.frame(
            IPTG_Conc1 = IPTG_conc1,
            IPTG_Conc2 = IPTG_conc2,
            t_statistic = t_statistic,
            SEM = se_diff,
            degrees_of_freedom = df,
            p_value = p_value,
            Bonferroni_adj = bonferroni_adj
          )
        }
      }
    }
  }

  do.call(rbind, results)
}



# Perform pairwise comparisons within each strain
ab_comparisons <- perform_pairwise_t_tests(combined_aba_stats)
eco_comparisons <- perform_pairwise_t_tests(combined_eco_stats)
```

##### Welch's t-test results
```{r, echo=FALSE}
print("A. baumannii stats")
print(ab_comparisons)
```

```{r, echo=FALSE}
print("E. coli stats")
print(eco_comparisons)
```




## Microscopy image analysis for plasmid expression (Fig S3)

Load all data tables (.csv for each sample) and sample names
```{r echo=FALSE}
mic_data <- fread("C:/Users/jentr/OneDrive - UW-Madison/lab things/data/Ab_plasmids/microscopy tifs/2024_08_19/20240819 microscopy fluorescent means.csv", header = TRUE)


mic_sample_map <- fread("C:/Users/jentr/OneDrive - UW-Madison/lab things/data/Ab_plasmids/microscopy tifs/2024_08_19/20240819 microscopy samples.csv", header = TRUE)

glimpse(mic_data)
```

```{r echo=FALSE}
print(mic_sample_map)
```

Convert data to long format and combine with map

```{r}
# Convert to long format
mic_long <- mic_data %>%
  pivot_longer(cols = everything(), names_to = "Sample_name", values_to = "fluor_int")

# Join with the second dataframe
merged_mic <- mic_long %>%
  left_join(mic_sample_map, by = "Sample_name")

# View the resulting dataframe
head(merged_mic)
```


Plot the data as Sina plots.

First for A. baumannii.
```{r, warning=FALSE}
merged_mic <- merged_mic %>%
  mutate(IPTG_conc = factor(IPTG_conc, levels = c("EV", "0", "62.5", "1000")))

aba_data <- merged_mic %>% filter(Strain == "Aba")

# Call fluorescence threshold (mean of EV + 1 SD)
mean_threshold <- aba_data %>% filter(IPTG_conc == "EV") %>%
  summarise(mean_threshold = mean(fluor_int, na.rm = TRUE) + 
              sd(fluor_int, na.rm = TRUE)) %>%
  pull(mean_threshold)

# Create the Sina plots
ggplot(aba_data, aes(x = IPTG_conc, y = fluor_int, fill = IPTG_conc)) +
  geom_sina(alpha = 0.25, size = 2.5) +
  geom_hline(yintercept = mean_threshold, linetype = "dashed", 
             color = "darkgreen", linewidth=1.25) +
  scale_y_log10(limits = c(0.05, 100000)) +
  theme_minimal() +
  labs(title = "A. baumannii single cell fluorescence intensity",
       x = "IPTG Concentration",
       y = "Raw Fluorescence Intensity Values") +
  theme(legend.position = "none")
```


Calculate percent of cells that are above the mean+SD threshold above background
```{r}
IPTG_percent <- aba_data %>%
  filter(!is.na(fluor_int)) %>%  # Exclude rows where fluor_int is NA
  group_by(IPTG_conc) %>%
  summarise(
    total = n(),
    above_threshold = sum(fluor_int > mean_threshold, na.rm = TRUE),
    percent_above_threshold = above_threshold / total * 100
  )

print(IPTG_percent)

```



Then for E. coli.
```{r, warning=FALSE}
eco_data <- merged_mic %>% filter(Strain == "Eco")

# Call fluorescence threshold (mean of EV + 1 SD)
mean_threshold <- eco_data %>% filter(IPTG_conc == "EV") %>%
  summarise(mean_threshold = mean(fluor_int, na.rm = TRUE) + 
              sd(fluor_int, na.rm = TRUE)) %>%
  pull(mean_threshold)

# Create the Sina plots
ggplot(eco_data, aes(x = IPTG_conc, y = fluor_int, fill = IPTG_conc)) +
  geom_sina(alpha = 0.25, size = 2.5) +
  geom_hline(yintercept = mean_threshold, linetype = "dashed", 
             color = "darkgreen", linewidth=1.25) +
  scale_y_log10(limits = c(0.05, 100000)) +
  theme_minimal() +
  labs(title = "E. coli single cell fluorescence intensity",
       x = "IPTG Concentration",
       y = "Raw Fluorescence Intensity Values") +
  theme(legend.position = "none")
```
Calculate percent of cells that are above the mean+SD threshold above background
```{r}
IPTG_percent <- eco_data %>%
  filter(!is.na(fluor_int)) %>%  # Exclude rows where fluor_int is NA
  group_by(IPTG_conc) %>%
  summarise(
    total = n(),
    above_threshold = sum(fluor_int > mean_threshold, na.rm = TRUE),
    percent_above_threshold = above_threshold / total * 100
  )

print(IPTG_percent)

```


And run statistics to determine differences (Welch's t-tests)

```{r}
pairwise_results <- pairwise.t.test(aba_data$fluor_int, aba_data$IPTG_conc, 
                                    p.adjust.method = "bonferroni",  # Bonferroni correction
                                    na.action = na.omit)  # Omit NAs
print("A. baumannii stats")
print(pairwise_results)
```
```{r}
pairwise_results <- pairwise.t.test(eco_data$fluor_int, eco_data$IPTG_conc, 
                                    p.adjust.method = "bonferroni",  # Bonferroni correction
                                    na.action = na.omit)  # Omit NAs
print("E. coli stats")
print(pairwise_results)
```